%-------------------------------------------------------------------------------
% content.tex — CV CONTENT ONLY
% A flat list of declared elements. No sections, columns, pages, or layout code:
% all structure lives in layout.tex and is assembled by \makecv (see main.tex).
% Declare elements with the macros documented in layout.tex.
%-------------------------------------------------------------------------------

%======================= HEADER =======================
\name{GIACOMO MARCIANI}
\role{SENIOR SOFTWARE ENGINEER}
\company{Amazon Web Services}
\headline{Building High Performance Computing in the Cloud}
\location[https://www.google.com/maps/place/Boston,+MA]{Boston, MA, United States}
\email{giacomo.marciani@gmail.com}
\linkedin{gmarciani}
\github{gmarciani}

%======================= INTRODUCTION =======================
\introduction{I am a Senior Software Engineer with 8+ years at AWS, specializing in High
Performance Computing in cloud environments and large-scale infrastructure automation.
I own the full lifecycle end-to-end: strategy, design, implementation,
DevOps, and on-call operations, with experience leading technical teams
and shaping product direction.\\
I have contributed to AWS ParallelCluster, NICE EnginFrame, Elastic Graphics,
and Elastic Inference. My background spans software engineering, distributed
systems and big data. Before AWS, I founded the first ACM Student Chapter in Rome,
co-founded a startup, and taught computer science at university. I also led research
projects that resulted in publications in scientific journals and presentations at
international conferences.\\
What makes me effective as an engineer and a leader is a relentless curiosity;
a drive to keep improving how I build, how I lead, and how I show up for my team.}

%======================= WORK EXPERIENCE =======================
\job{Sr Software Engineer}{Amazon Web Services}{2021-Present}
{Technical leader of AWS ParallelCluster. My responsibilities span the software
lifecycle end-to-end: strategy, design, implementation, observability, on-call
incident response, mentoring, and hiring.}

\job{Software Engineer II}{Amazon Web Services}{2019-2021}
{Worked on Elastic Graphics, Elastic Inference, and NICE EnginFrame. On EnginFrame,
I contributed mainly to the build system, CI/CD automation, and integration work to
enable the portal to interact with both on-premises and cloud-based clusters. Starting
in 2021, I took on AWS ParallelCluster, where I became the team's de facto technical
leader. My key contributions include Multi-User environment, Dynamic File System
Mounting, support for GPU-intensive instance types, enabling ParallelCluster in US
isolated regions, Terraform integration, ParallelCluster UI, serving as Security
Guardian, driving roadmap and releases, and resolving high-severity escalations for
key customers. I also led the complex shutdown of Elastic Graphics and Elastic
Inference, navigating deep integrations with EC2/ECS.}

\job{Software Engineer I}{Amazon Web Services}{2018-2019}
{Worked on Elastic Graphics and Elastic Inference, which attached cost-effective
GPU acceleration to EC2 instances and other AWS services. I contributed to the
launch of Elastic Inference at Re:Invent 2018 and owned the metering and billing
component for compute resources.}

\job{R\&D Intern}{Accenture}{2017}
{Designed a solution for the elastic auto-scaling of containerized applications
using Reinforcement Learning in Kubernetes.}

\job{Chairman, Founder}{ACM Rome Tor Vergata}{2016-2018}
{Founded the ACM Student Chapter at University of Rome Tor Vergata, the first in
the city. Led the team and fostered connections between academia and industry.}

\job{Teaching Assistant}{University of Rome Tor Vergata}{2016-2018}
{Tutored undergraduate students and graded exams for Foundations of Computer
Science and Algorithms Engineering.}

\job{Mathematics Tutor}{Freelance}{2011-2017}
{Gave private math lessons to secondary and high school students. Several of my
students went on to study engineering.}

\job{Management Engineer, Co-Founder}{Sbaam.com, e-Commerce}{2011-2012}
{Co-founded the startup, making it a finalist in the InnovActionLab startups
contest. Responsible for business plan development and investor pitches.}

%======================= MAIN PROJECTS =======================
\project[https://aws.amazon.com/hpc/parallelcluster/]{AWS ParallelCluster and ParallelCluster UI}{Amazon Web Services, 2021-Present | HPC}
{Open-source platform to automate deployment and lifecycle management of large-scale
HPC clusters on AWS, with millions of dollars in revenue and major enterprise customers.}

\project[https://www.ni-sp.com/products/ef-hpc-portal/]{NICE EnginFrame}{Amazon Web Services, 2021 | HPC}
{Web application to submit HPC workloads to both on-premises and AWS clusters.}

\project[https://aws.amazon.com/about-aws/whats-new/2018/11/introducing-amazon-elastic-inference/]{AWS Elastic Graphics and Elastic Inference}{Amazon Web Services, 2018-2023 | Web Service}
{Services to attach GPU-powered acceleration for OpenGL and ML workloads to EC2,
ECS and Sagemaker to reduce the cost of running graphics and machine learning
applications.}

\project[https://ieeexplore.ieee.org/abstract/document/8113784]{Crime Graph}{Big Data, IEEE FiCloud, 2017 | Java, Kafka, Flink, Neo4j}
{Data stream processing application for real-time analysis of evolving criminal
networks, uncovering hidden and latent relationships.}

% \project[https://github.com/gmarciani]{NNML-HPC}{High Performance Computing, 2017 | C, CUDA}
% {Deep neural network for handwritten digit recognition, implemented in CUDA/C
% with OpenMP.}

\project[https://github.com/gmarciani/pydes]{PyDES}{Performance Modeling, 2019 | Python}
{Discrete-event simulation suite for performance modeling of cloud computing systems,
featuring batch-means analysis, Markov chain solvers, and multi-stream pseudo-random
number generation.}

\project[https://github.com/acm-utv/ontoqa]{OntoQA}{Artificial Intelligence, 2017 | NLP, Java, Protege}
{Question\&Answer web application leveraging ontology-driven natural language
processing.}

\project[https://github.com/gmarciani/godot-android]{Godot}{App for Transports, 2013 | Java, Android}
{Mobile app to address issues caused by illegal parking, leveraging QR codes and
NFC.}
\sectionfootnote{See more on my \href{https://github.com/gmarciani}{GitHub}}

%======================= PUBLICATIONS =======================
\publication[https://ieeexplore.ieee.org/abstract/document/8113784]{A data streaming approach to link mining in criminal networks}{5th IEEE International Conference on Future IoT (FiCloud 2017)}
\publication[https://dl.acm.org/doi/abs/10.1145/2933267.2933517]{Real-time analysis of social networks leveraging Flink}{10th ACM International Conference on Distributed and Event-Based Systems (DEBS 2016)}
\publication[https://arxiv.org/abs/1611.01866]{Marciani Normal Form of context-free grammars}{Arxiv (2016)}

%======================= EDUCATION =======================
\education{MSc Software Engineering and Data Science}{University of Rome Tor Vergata, 2015-2018}{Grad Avg 29/30 - GPA 4/4 *}
\education{BSc Software Engineering}{University of Rome Tor Vergata, 2011-2015}{Grad 110/110 cum laude}
\education{BSc Management Engineering}{University of Rome Tor Vergata, 2009-2011}{Interrupted}
% \education{Cambridge Summer School}{Girton College, Cambridge, UK, 2005-2006}{Grad English Proficient}
% \education{Classical License}{Liceo T. Tasso, Rome, IT, 2004-2009}{Grad 77/100}

%======================= MAIN COURSES =======================
% \course{Foundation of Computer Science}{30L}
% \course{Software Engineering}{30L}
% \course{Internet Engineering}{30L}
% \course{Web Programming}{30L}
% \course{Web Algorithms}{30L}
% \course{Big Data Architectures}{30L}
% \course{Artificial Intelligence}{30L}
% \course{Distributed Systems}{30L}
% \course{Computer Security}{30L}
% \course{Computer Architectures}{30}
% \course{Optimization Methods}{29}
% \course{Mobile Programming}{29}
% \course{Algorithms Engineering}{29}
% \course{Multi Core Programming}{29}
% \course{Performance Modeling}{29}
% \course{Operations Research}{27}
% \course{Automata and Languages}{26}
% \course{Economics}{30}
% \course{ICT Project Analysis}{30}

%======================= HONOURS =======================
\honour{Lecturer at Master CESMA}{University of Rome Tor Vergata, 2016}
\honour{Finalist at ACM DEBS Grand Challenge}{10th ACM International Conference on Distributed Event Based Systems, 2016}
\honour{Speaker at InnovAction Lab Grand Challenge}{LVenture Groups, 2013}
\honour{Best Student in Computer Architectures}{University of Rome Tor Vergata, 2012}
\honour{Alumno of InnovAction Lab}{LVenture Groups, 2011}
\honour{Counselor of Engineering}{University of Rome Tor Vergata, 2010}

%======================= MAIN SKILLS =======================
\skill{Cloud}{Amazon Web Services, Digital Ocean, Railway}
\skill{HPC}{ParallelCluster, EnginFrame, Slurm, Nvidia GPUs, Cuda, MPI, Lustre, EFA}
\skill{Big Data}{Flink, Hadoop Distributed File System, Hadoop MapReduce, Pig, Hive, Kafka, Flume, Elasticsearch, Kibana, Grafana, Kubernetes}
\skill{Web}{Spring, Nginx, Tomcat, React, NodeJS, HTML, CSS, Sass, Postman, Hugo, Wordpress}
\skill{Automation}{CloudFormation, Terraform, Chef}
\skill{Data}{SQL, NoSQL, Graph, Caching}
\skill{Languages}{Python (primary), Java, Kotlin, Ruby, Javascript, C, Cuda, Bash, HTML, OWL, SPARQL, AMPL, MIPS, Latex}
\skill{Platforms}{Linux, OSX, Docker}
\skill{Applied Math}{Matlab, AMPL, IBM Cplex}
\skill{Process}{Agile, SCRUM}

%======================= LANGUAGES =======================
\lang{Italian}{Native}
\lang{English}{Proficient}

%======================= FOOTER =======================
% \footer{Maximum is 30/30. L stands for cum laude. Grade 27 is equivalent to A.
% * Thesis pending. I hereby authorize to use and process my personal details contained
% in this document (Italian Decree 196 30/06/2003).}
